---
title: "TMDB Box Office Prediction"
output:
  word_document: default
  html_document: default
---

# As of 2018, global box office is worth $41.7 billion dollars. But what makes a movie a top hit? What are the key drivers of revenue? This project will use the data from the movie database Kaggle competition to explore the key drivers of successful movies and develop machine learning models to predict their box office revenues

# OUTLINE
# Exploratory Data Analysis & Data Cleaning
# Feature Enginnering (create 5 new features)
# Preprocess
# Build Models (4 linear regression and 3 decision tree regression models)
# Model Comparison
# Conclusion


# Set up
```{r}
# load libraries
library(tidyverse)
library(GGally)
library(caret)
library(lubridate) # date
library(naniar) # provide plots to explore missing data
library(randomForest)
library(rpart.plot)

# packages for generating a word map
library(tm) # test mining
library(SnowballC) # text stemming
library(wordcloud)
library(XML)
library(RColorBrewer) # for color palettes

# make sure to get identical result each time running the code
set.seed(1234)
```

# Read TMDB dataset
```{r}
df_train = read_csv("train.csv")
df_test = read_csv("test.csv")

# combine the train and test dataset
data = bind_rows(df_train, df_test)

df_full = data
```


# Exploratory Data Analysis
```{r}
# take a high_level view of the dataset
glimpse(df_full)   # there are 7,398 observations and 23 variables; 3,000 observations are for training
summary(df_full)

# Let's look at how the movie industry has developed over the years
# convert release_date column to date format
df_full$release_date = as.Date(df_full$release_date, format = "%m/%d/%y")

# extract year from the date
df_full$year = as.numeric(format(df_full$release_date, "%Y"))

# extract month
df_full$month = as.numeric(format(df_full$release_date, "%m"))

# rename the month
levels(df_full$month)[levels(df_full$month)==1] = "Jan"
levels(df_full$month)[levels(df_full$month)==2] = "Feb"
levels(df_full$month)[levels(df_full$month)==3] = "Mar"
levels(df_full$month)[levels(df_full$month)==4] = "Apr"
levels(df_full$month)[levels(df_full$month)==5] = "May"
levels(df_full$month)[levels(df_full$month)==6] = "Jun"
levels(df_full$month)[levels(df_full$month)==7] = "Jul"
levels(df_full$month)[levels(df_full$month)==8] = "Aug"
levels(df_full$month)[levels(df_full$month)==9] = "Sep"
levels(df_full$month)[levels(df_full$month)==10] = "Oct"
levels(df_full$month)[levels(df_full$month)==11] = "Nov"
levels(df_full$month)[levels(df_full$month)==12] = "Dec"

# plot the number of movies over years
ggplot(df_full, aes(x = year)) +
         geom_bar(width = 1, fill="steelblue") +
         xlim(1970, 2018) # I set limit for x axis because some release years are in the future

# plot number of movies by month
ggplot(df_full, aes(x = month)) +
         geom_bar(fill="steelblue")   # looks like summer is peak season

# plot the distribution of revenue
ggplot(df_full[1:3000, ], aes(x = revenue)) +
      geom_bar(width = 0.1, fill="steelblue") +
      scale_x_log10() + 
      geom_vline(aes(xintercept = mean(revenue)), color = "red", linetype = "dashed")

# plot word map for movie titles
title_corpus = Corpus(VectorSource(df_full$title))
title_clean = tm_map(title_corpus, PlainTextDocument) # convert to plain text
title_clean = tm_map(title_clean, tolower) # convert to lowercases
title_clean = tm_map(title_clean,removeWords,stopwords("english")) # remove stopwprds
wordcloud(words = title_clean, min.freq = 3, max.words = 100, colors = brewer.pal(8, "Dark2"))

# plot word cloud for movie taglines
tagline_corpus = Corpus(VectorSource(df_full$tagline))
tagline_clean = tm_map(tagline_corpus, PlainTextDocument)
tagline_clean = tm_map(tagline_clean, tolower) 
tagline_clean = tm_map(tagline_clean,removeWords,stopwords("english")) 
wordcloud(words = tagline_clean, min.freq = 3, max.words = 100, colors = brewer.pal(8, "Dark2"))

# use ggpairs to explore possible corelations
ggpairs(select(df_full[1:3000, ], revenue, popularity, budget, runtime)) 
# budget has a positive high corelation with revenue (0.753)

# get summary of the missing value in the datasets
colSums(is.na(df_full))

# plot the percetage of missing values
gg_miss_var(df_full, show_pct = TRUE) +
           labs(y = "Percetage of Missing Values for TMDB Datasets")
```

# Feature Engineering
# I noticed that some columns like belongs_to_collection and production companies are quite messy as the core information is mixed with irrelevant symbols and ids
# So I will try to clean them first (I extract keywords by regular expression)

1. belongs_to_collection
```{r}
# Extract the name of collections
df_full$belongs_to_collection = str_extract(df_full$belongs_to_collection, 
                               pattern = "(?<=name\\'\\:\\s{1}\\').+(?=\\'\\,\\s{1}\\'poster)")

df_full$belongs_to_collection = as.factor(df_full$belongs_to_collection)

# check the top collections by average revenue
df_full[1:3000, ] %>%
  group_by(belongs_to_collection) %>%
  summarise(avg_rev = mean(revenue)) %>%
  arrange(desc(avg_rev)) %>%
  filter(!is.na(belongs_to_collection))

# Hypothesis: revenue for movies in a collection are higher than those not in a collection
# create a collection column to classify whether movies are in a collection or not
# Feature Engineering - Movie collections
df_full$collection = 0 

for(i in 1:length(df_full$belongs_to_collection)) {
  if(is.na(df_full$belongs_to_collection[i])) {
   df_full$collection[i] = "no"
  }
  else {
    df_full$collection[i] = "yes"
  }
}

# plot the revenue for movies in or not in an collection
ggplot(data = df_full[1:3000, ], aes(x = collection, y = revenue, fill = collection)) +
              geom_boxplot(alpha = 0.3)

```

2. production countries
```{r}
# 2. production countries 
df_full$production_countries = str_extract(df_full$production_countries, pattern = "[:upper:]+")

# check the top production countries by average revenue
df_full[1:3000, ] %>%
  group_by(production_countries) %>%
  summarise(movies_count = n()) %>%
  arrange(desc(movies_count)) %>%
  filter(!is.na(production_countries)) %>%
  head(10)

# Hypothesis: top production countries have higher box office revenue
# Feature Engineering - top production countries

df_full$top_production_countries = 0 

for(i in 1:length(df_full$production_countries)) {
  if(df_full$production_countries[i] %in% c("US", "GB", "FR") ) {
   df_full$top_production_countries[i] = "yes"
  }
  else  {
    df_full$top_production_countries[i] = "no"
  }
}
# plot
ggplot(df_full[1:3000, ], aes(x = top_production_countries, y = revenue, fill = top_production_countries)) +
             geom_boxplot(alpha = 0.3)

```

3. production companies
```{r}
# 3. production companies
df_full$production_companies = gsub('(^\\[\\{\'name\'\\:\\s\'|\'\\,\\s\'id.*)', '',
                                 df_full$production_companies)

# check the top production companies by average revenue
df_full[1:3000, ] %>%
  group_by(production_companies) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  filter(!is.na(production_companies))

# Hypothesis: large production companies make more money
# Feature Engineering - top production companies

df_full$top_production_companies = 0 

for(i in 1:length(df_full$production_companies)) {
  if(df_full$production_companies[i] %in% c("Universal Pictures", 
                                            "Paramount Pictures",
                                            "Twentieth Century Fox Film Corporation",
                                            "Columbia Pictures",
                                            "Warner Bros.",
                                            "New Line Cinema",
                                            "Walt Disney Pictures") ) {
   df_full$top_production_companies[i] = "yes"
  }
  else  {
    df_full$top_production_companies[i] = "no"
  }
}
# plot
ggplot(data = df_full[1:3000, ], aes(x =top_production_companies, y = revenue, fill = top_production_companies )) +
             geom_boxplot(alpha = 0.3)

```

4. Languages
```{r}
# count the top movie languages
df_full[1:3000,] %>%
   group_by(original_language) %>%
   summarise(count = n()) %>%
   arrange(desc(count))

# most movies have enghlish as their original languages
# Hypothesis: english speaking movies will generate more revenue
# feature engineering: eng_movie
df_full$eng_movie = 0 

for(i in 1:length(df_full$original_language)) {
  if(df_full$original_language[i] == "en") {
   df_full$eng_movie[i] = "yes"
  }
  else {
    df_full$eng_movie[i] = "no"
  }
}

# plot the revenue for english movies and all other movies
ggplot(df_full[1:3000, ], aes(x = eng_movie, y = revenue, fill = eng_movie)) +
             geom_boxplot(alpha = 0.3) +
             xlab("Homepage presence")
```

5. Feature Engineering - Homepage
```{r}
# Hypothesis: box office revenue will be higher if they build a homepage

df_full$Hpage_check = 0 

for(i in 1:length(df_full$homepage)) {
  if(is.na(df_full$homepage[i])) {
   df_full$Hpage_check[i] = "no"
  }
  else {
    df_full$Hpage_check[i] = "yes"
  }
}

# plot the revenue for movies in or not in an collection
ggplot(data = df_full[1:3000, ], aes(x = Hpage_check, y = revenue, fill = Hpage_check)) +
              geom_boxplot(alpha = 0.3) + 
              xlab("Homepage presence")
        
```

# Split the datasets into training and testing
```{r}
# select variables for modeling
df_full_final = select(df_full, revenue, budget, popularity, runtime, year, collection, top_production_companies,
                        top_production_countries, eng_movie, Hpage_check)

# check the missing value of the data
colSums(is.na(df_full_final)) 

# replace missing values by their mean, except for revenue which is our outcome variable
df_full_final$runtime[is.na(df_full_final$runtime)] = mean(df_full_final$runtime, na.rm = TRUE)
df_full_final$year[is.na(df_full_final$year)] = mean(df_full_final$year, na.rm = TRUE)

# check near zero values
nearZeroVar(df_full_final, saveMetrics = TRUE) # no near zero values

# convert categorical variables to binary features for linear regression
df_full_linear = df_full_final

df_full_linear = df_full_linear %>% mutate_if(is.character, as.factor)

collection = model.matrix(~collection - 1, data = df_full_linear)
top_production_companies = model.matrix(~top_production_companies - 1, data = df_full_linear)
eng_movie = model.matrix(~eng_movie - 1, data = df_full_linear)
nrow(top_production_companies)
Hpage_check = model.matrix(~Hpage_check - 1, data = df_full_linear)

df_full_linear = cbind(df_full_linear, collection, top_production_companies, eng_movie, Hpage_check)
df_full_linear = select(df_full_linear, -collection, -top_production_companies, -eng_movie, -Hpage_check)

```

# Build Models

1. Linear Regresssion
b. forward regression
```{r}
# split the data
df_train1 = df_full_linear[1:3000, ]
df_test1 = df_full_linear[3001:7398, ]

forward_model = train(revenue ~ ., data = df_train1,
                      method = "leapForward", tuneGrid = expand.grid(nvmax = 1:10), 
                      trControl = trainControl(method = "cv", number = 10))

forward_model
plot(forward_model)
# extract the final model
plot(forward_model$finalModel, scale = "adjr2")
```

c. ridge regression
Address the problem of parameter istability caused by highly correlated features
shrink the unstable coefficients by imposing an additional penalty parameter on the errors of an ordinary least squares regression
penalize coefficients that are either too big or too small
```{r}
ridge_model = train(revenue ~., data = df_train1,
                    method = "ridge", tuneGrid = expand.grid(lambda = seq(0, 1, 0.05)),
                    trControl = trainControl(method = "cv", number = 10))

ridge_model
plot(ridge_model)

# extract the final model
plot(ridge_model$finalModel)
```

d. lasso model
```{r}
lasso_model = train(revenue ~., data = df_train1,
                    method = "lasso", tuneGrid = expand.grid(fraction = seq(0, 1, 0.5)),
                    trControl = trainControl(method = "cv", number = 10))
lasso_model

plot(lasso_model)

# extract the final model
plot(lasso_model$finalModel)

# get the model coefficients
lasso_coefs = predict(lasso_model$finalModel, type = "coef")
```
d. simple linear regression model
```{r}
fit = train(revenue ~ collectionyes + popularity + budget + eng_movieyes, data = df_train1, 
            method = "lm",trControl = trainControl(method = "cv", number = 10), metric = "RMSE")

fit

# average parameter values after a 10-fold CV
fit$finalModel
summary(fit)
```

```{r}
# predict revenue on the test data
prediction = predict(fit, newdata = df_test1)
solution = data.frame(id = 3001:7398, revenue = prediction)

# write solution to file and submit to Kaggle website
# write.csv(solution, file = "linear_prediction.csv", row.names = FALSE)

# linear model has the best average RMSE
# Kaggle evaluated based on RMSD (Root mean square deviation) 7.36271

```

Linear regression model comparison
```{r}
compare = resamples(list(forward_selection = forward_model, 
                         ridge_selection = ridge_model, lasso_selection = lasso_model, 
                         linear = fit))

# compare RMSE and R-squared
summary(compare)

# plot the result
dotplot(compare)
```
One tree model
```{r}
training_set = select(df_train1, -revenue)
tree_model = train(y = df_train1$revenue, x = training_set, method = "rpart")

tree_model

rpart.plot(tree_model$finalModel)

plot(varImp(tree_model))
```

Bagging model
```{r}
bagged_model = train(y = df_train1$revenue, x = training_set, method = "treebag")
bagged_model
plot(varImp(bagged_model))
```

Random forest regression model
```{r}
rf_model <- randomForest(revenue ~ ., data = df_train1, ntree = 300, replace = TRUE, importance = TRUE)

rf_model
plot(rf_model)
```

```{r}
# predict revenue on the test data
prediction1 = predict(rf_model, newdata = df_test1)
solution1 = data.frame(id = 3001:7398, revenue = prediction1)

# write solution to file and submit to Kaggle website
write.csv(solution1, file = "tree_prediction.csv", row.names = FALSE)

# Kaggle score is 2.68194
```



# Conclusion
```{r}
#If you are an entrepreneur who are looking for investment opportunities in the movie industry, you should consider movies have:
# 1.	Big budgets
# 2.	High popularity of content
# 3.	is part of a popular collection
# 4.	Not short films
# 5.	From big production companies and countries
# 6.	Have a homepage

# Next Step of the project to predict accuracy:
# 1. adjust the movie budget and revenue for inflation 
# 2. cleaning cast and crew columns to see how famous directors and actors influence the revenue

```

